Oregon 2013 Medication for Asthma Inhalers

### 2017-06-09 r_scripts/analysis/oregon_medication.Rmd
add four disease category -need modify explanation

```{r library, include=FALSE, echo=FALSE}
library(dplyr)
library(data.table)
library(readxl) 
library(readr)
library(survival)
library(ggplot2)
library(htmlTable)
```

@@

or_asthma_miss <- or_asthma %>%
  filter(dx1=="*NULL*" &
         dx2=="*NULL*" &
         dx3=="*NULL*" &
         dx4=="*NULL*" &
         dx5=="*NULL*" &
         dx6=="*NULL*" &
         dx7=="*NULL*" &
         dx8=="*NULL*" &
         dx9=="*NULL*" &
         dx10=="*NULL*" &
         dx11=="" &
         dx12=="" &
         dx13=="") %>%
  filter(ndc=="" |
         ndc=="*NULL*") # 20269 missing in total data set

or_copd_miss <- or_copd %>%
  filter(dx1=="*NULL*" &
         dx2=="*NULL*" &
         dx3=="*NULL*" &
         dx4=="*NULL*" &
         dx5=="*NULL*" &
         dx6=="*NULL*" &
         dx7=="*NULL*" &
         dx8=="*NULL*" &
         dx9=="*NULL*" &
         dx10=="*NULL*" &
         dx11=="" &
         dx12=="" &
         dx13=="") %>%
  filter(ndc=="" |
         ndc=="*NULL*") # 20269 missing in total data set

or_both_miss <- or_both %>%
  filter(dx1=="*NULL*" &
         dx2=="*NULL*" &
         dx3=="*NULL*" &
         dx4=="*NULL*" &
         dx5=="*NULL*" &
         dx6=="*NULL*" &
         dx7=="*NULL*" &
         dx8=="*NULL*" &
         dx9=="*NULL*" &
         dx10=="*NULL*" &
         dx11=="" &
         dx12=="" &
         dx13=="") %>%
  filter(ndc=="" |
         ndc=="*NULL*") # 20269 missing in total data set

or_other_miss <- or_other %>%
  filter(dx1=="*NULL*" &
         dx2=="*NULL*" &
         dx3=="*NULL*" &
         dx4=="*NULL*" &
         dx5=="*NULL*" &
         dx6=="*NULL*" &
         dx7=="*NULL*" &
         dx8=="*NULL*" &
         dx9=="*NULL*" &
         dx10=="*NULL*" &
         dx11=="" &
         dx12=="" &
         dx13=="") %>%
  filter(ndc=="" |
         ndc=="*NULL*") # 20269 missing in total data set

@@

@@

### NDC claims with index

@@

>>>>

```{r library, include=FALSE, echo=FALSE}
library(dplyr)
library(data.table)
library(readxl) 
library(readr)
library(survival)
library(ggplot2)
library(htmlTable)
library(knitr)
```

@@

or_asthma_miss <- or_asthma %>%
  filter(dx1=="*NULL*" &
         dx2=="*NULL*" &
         dx3=="*NULL*" &
         dx4=="*NULL*" &
         dx5=="*NULL*" &
         dx6=="*NULL*" &
         dx7=="*NULL*" &
         dx8=="*NULL*" &
         dx9=="*NULL*" &
         dx10=="*NULL*" &
         dx11=="" &
         dx12=="" &
         dx13=="") %>%
  filter(ndc=="" |
         ndc=="*NULL*") # 89482 claims missing in asthma data set

or_copd_miss <- or_copd %>%
  filter(dx1=="*NULL*" &
         dx2=="*NULL*" &
         dx3=="*NULL*" &
         dx4=="*NULL*" &
         dx5=="*NULL*" &
         dx6=="*NULL*" &
         dx7=="*NULL*" &
         dx8=="*NULL*" &
         dx9=="*NULL*" &
         dx10=="*NULL*" &
         dx11=="" &
         dx12=="" &
         dx13=="") %>%
  filter(ndc=="" |
         ndc=="*NULL*") # 18941 claims missing in copd data set

or_both_miss <- or_both %>%
  filter(dx1=="*NULL*" &
         dx2=="*NULL*" &
         dx3=="*NULL*" &
         dx4=="*NULL*" &
         dx5=="*NULL*" &
         dx6=="*NULL*" &
         dx7=="*NULL*" &
         dx8=="*NULL*" &
         dx9=="*NULL*" &
         dx10=="*NULL*" &
         dx11=="" &
         dx12=="" &
         dx13=="") %>%
  filter(ndc=="" |
         ndc=="*NULL*") # 17963 claims missing in both data set

or_other_miss <- or_other %>%
  filter(dx1=="*NULL*" &
         dx2=="*NULL*" &
         dx3=="*NULL*" &
         dx4=="*NULL*" &
         dx5=="*NULL*" &
         dx6=="*NULL*" &
         dx7=="*NULL*" &
         dx8=="*NULL*" &
         dx9=="*NULL*" &
         dx10=="*NULL*" &
         dx11=="" &
         dx12=="" &
         dx13=="") %>%
  filter(ndc=="" |
         ndc=="*NULL*") # 46971 claims missing in other data set (5848 personkey)

or_other_not_miss <- or_other %>%
  filter(dx1!="*NULL*" |
         dx2!="*NULL*" |
         dx3!="*NULL*" |
         dx4!="*NULL*" |
         dx5!="*NULL*" |
         dx6!="*NULL*" |
         dx7!="*NULL*" |
         dx8!="*NULL*" |
         dx9!="*NULL*" |
         dx10!="*NULL*" |
         dx11!="" |
         dx12!="" |
         dx13!="" |
         ndc!="" |
         ndc!="*NULL*") # 20269 missing in total data set

length(unique(or_other_not_miss$personkey)) # [1] 27275

@@

  print(print_plot)
  # save figure
  ggsave("../plot_new/ndc_overall_three.pdf", plot = print_plot, 
       width = 12, height = 8, units = "in")

@@

## NDC claims with index

Check the claims in each category (Asthma, COPD, Both or Other).

@@

```{r ndc descriptive table, echo = F, message = F, warning = F, results='asis'}
read_path6 <- paste0("../data_new/medication/oregon_ndc_time_strat_casecrossover.csv")
beta_ndc <- read_csv(read_path6)

# join personkey with index
oregon_beta_index_df <- oregon_beta_df %>%
  select(personkey, index)

beta_ndc_with_index <- beta_ndc %>%
  left_join(oregon_beta_index_df, by="personkey")

# dataframe to loop through
beta_ndc_with_index_df <- data.frame(beta_ndc_with_index)

beta_asthma <- beta_ndc_with_index_df %>%
  filter(index=="asthma")

beta_copd <- beta_ndc_with_index_df %>%
  filter(index=="copd")

beta_both <- beta_ndc_with_index_df %>%
  filter(index=="both")

beta_other <- beta_ndc_with_index_df %>%
  filter(index=="other")

df_list <- list(beta_asthma, beta_copd, beta_both, beta_other)

outcome_list <- c("Asthma", "COPD", "Both", "Other")

# create an empty list to row bind dataframes together
datalist <- list()

# outcome name
outcome_name <- c("Asthma", "COPD", "Both", "Other")

for(i in 1:length(df_list)){
  # dataframe to loop through
  df_to_loop <- data.frame(df_list[i])
  # indication of column
  outcome <- colnames(df_to_loop[75])
  # outcome name
  outcome_name <- outcome_list[i]

  # dataframe for analysis creation
  # bind columns back together 
  df_analysis <- df_to_loop %>%
    filter(!is.na(wrf_smk_pm_zip)) %>% 
    # only look at outcomes
    filter(outcome == 1) %>%
    # add another row that makes sure there is a person <15 in the dataframe
    # tricking xtabs to produce a 0 cell for the outcome for age <15
    add_row(outcome = 0, age_ind = 0)

  # combine previous values in dataframe that has all outcome/methods comparisons
  datalist[[i]] <- point_est_df

} # end of loop

# combine each outcome dataframe itteration in to a big dataset
combined_point_est_df <- bind_rows(datalist) %>% 
  # find proportions/percents for each strata in a row
  mutate(age_15_pr = as.character(round((as.numeric(age_15)/as.numeric(n))*100,1)),
         age_15_65_pr = as.character(round((as.numeric(age_15_65)/as.numeric(n))*100,1)),
         age_65_pr = as.character(round((as.numeric(age_65)/as.numeric(n))*100,1)),
         female_pr = as.character(round((as.numeric(female)/as.numeric(n))*100,1)),
         male_pr = as.character(round((as.numeric(male)/as.numeric(n))*100,1))) %>% 
  select(outcome, n, age_15, age_15_pr, age_15_65, age_15_65_pr, age_65,
         age_65_pr, female, female_pr, male, male_pr)

#str(combined_point_est_df)

tab <- htmlTable(txtRound(combined_point_est_df, digits = 1), 
           caption = "Number of cases for each outcome observed from May 1st to September 30st, 2013",
           # column headers
           header = c("Outcome", "Cases n", "Less than 15", "(%)", "15 to 65", "(%)", "Greater than 65",
                      "(%)", "Female", "(%)", "Male", "(%)"),
           # column spanner
           cgroup = c("","Age Category", "Sex"), 
           n.cgroup = c(2, 6, 4),
           padding.rgroup = "&nbsp;&nbsp;",
           css.cell = "padding-left: 0.5em; padding-right: .5em;", # cell space
           align = "llccccc" # column alignment,
            ) # end table

```{r ndc overall try, warning =F, echo = F, results='asis'}
df_list <- list(beta_asthma, beta_copd, beta_both, beta_other)

outcome_list <- c("Asthma", "COPD", "Both", "Other")

# create an empty list to row bind dataframes together
datalist <- list()

# dataframe list
method_list <- c('WRF-Chem Smoke', 'Kriging Smoke', 'Geo-Weighted Smoke')

# outcome name
outcome_name <- c("Asthma", "COPD", "Both", "Other")

# data wrangling ----
# Producing conditional logit model estimates loop 
for(i in 1:length(df_list)){
  # dataframe to loop through
  df_to_loop <- data.frame(df_list[i])
  # indication of column
  outcome <- colnames(df_to_loop[75])
  # outcome name
  outcome_name <- outcome_list[i]

  # extract covariates from dataframe
  covariates_df <- df_to_loop[, c(1:26, 74:84)]

  # extract pm values and divide by 10 and ordered
  #which(colnames(df_to_loop)=="global_smk_pm_zip") # code to find column numbers
  pm_estimates_df <- df_to_loop[, c(87, 92, 91, 93)]/10  # create 10 unit increases

  # dataframe for analysis creation
  # bind columns back together 
  df_analysis <- cbind(covariates_df, pm_estimates_df) %>% 
    # remove missing pm values
    filter(!is.na(wrf_smk_pm_zip)) %>%

    # join in cdc met data
    # full_join(daily_zip_met, by = c("ZIPCODE", "date")) %>% 
    # the following code makes sure that the counterfactual values retained are 
    # symetric in that number of obs before = number of obs after
    mutate(obs_diff_admission = (fromdate - date)/7) 
  # dataframe is already for the entire fire season, so I don't need to subset anymore

  # empty df for table
  table_df <- data.frame()

  # empty matrix
  point_estimates <- matrix(nrow = 3, ncol = 9, byrow = T)

  # fill in the outcome name for the dataframe before the loop
  point_estimates[, 1] <- outcome_name

  # second loop to run a model for each pm estimation method

  for(j in 38:40){

  # combine previous values in dataframe that has all outcome/methods comparisons
  datalist[[i]] <- point_est_df

} # end of loop

# combine each outcome dataframe itteration in to a big dataset
combined_point_est_df <- bind_rows(datalist)

# subset columns I want to put in to the table
table_df <- combined_point_est_df %>% select(2, 3:7)

tab <- htmlTable(txtRound(table_df, digits = 3, 1:3), 
                 caption = "Association between a 10 ug/m^3 in PM2.5 and Health Outcomes",
                 # row group by outcome
                 rgroup = outcome_list,
                 n.rgroup = c(rep(3, 4)), # 4 rows for each method for each outcome
                 # column headers
                 header = c("Method", "Obs.", "Events",
                            "OR&dagger;", "Lower", "Upper"),
                 # column spanner
                 cgroup = c("", "95% CI"), 
                 n.cgroup = c(4, 2),
                 padding.rgroup = "&nbsp;&nbsp;",
                 css.cell = "padding-left: 0.5em; padding-right: .5em;", # cell space
                 align = "llccccc", # column alignment,
                 tfoot="&dagger; Adjusted for CDC temperature, relatively humidity, wind speed, and precipitation; accounting for subject. Time-stratified: referent periods matched to events on same day of week within July to October fire season."
) # end table

# ggplot of odds ratios, facet wrapped by outcomes -----
# convert variables from character to either numeric or factor
# factor preserves the order of the variable
combined_point_est_df$outcome <- factor(combined_point_est_df$outcome, 
                                        levels = unique(combined_point_est_df$outcome))

combined_point_est_df$pm_method <- factor(combined_point_est_df$pm_method, 
                                          levels = unique(combined_point_est_df$pm_method))

combined_point_est_df$n <- as.numeric(combined_point_est_df$n)
combined_point_est_df$n_events <- as.numeric(combined_point_est_df$n_events)
combined_point_est_df$odds_ratio <- as.numeric(combined_point_est_df$odds_ratio)
combined_point_est_df$lower95 <- as.numeric(combined_point_est_df$lower95)
combined_point_est_df$upper95 <- as.numeric(combined_point_est_df$upper95)
combined_point_est_df$se <- as.numeric(combined_point_est_df$se)
combined_point_est_df$p_val <- as.numeric(combined_point_est_df$p_val)

# subset out global method
combined_point_est_df <- combined_point_est_df

## ggplot
print_plot <- ggplot(combined_point_est_df, 
                     aes(x = pm_method, y = odds_ratio, colour = pm_method)) +
  # custom color 
  scale_color_manual(name = "Smoke-Estimation Method", 
                     values = c("red", "blue", "#32115C"),
                     guide = guide_legend(title.position = "top", title.hjust = 0.5)) +
  geom_point() + #geom_text(vjust = 0, nudge_x = 0.3) +
  geom_errorbar(aes(ymin=lower95, ymax=upper95), width = 0.2) +
  facet_wrap(~outcome, nrow = 3) +
  geom_hline(yintercept = 1, linetype=2) +
  #    ggtitle('Association Between PM2.5 \n from Wildfire Smoke on Hospitalizations') +
  ylab(expression(paste("Odds Ratio for 10µg/m"^3, " Increase in PM"[2.5]))) +
  #xlab('Smoke Estimation Method') +
  # plot theme
  theme(panel.background = element_rect(fill = 'white', colour = 'black'),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        # strip element
        strip.background = element_rect(colour=NA, fill=NA),
        panel.border = element_rect(fill = NA, color = "black"),
        # facet text size
        #strip.text = element_text(size = 8),
        # axis element
        axis.text.x = element_blank(),
        axis.title.x = element_blank(),
        axis.ticks.x = element_blank(),
        axis.title.y = element_text(angle = 90),
        # legend elements
        legend.position = "bottom")
#legend.text = element_text(size = 8))

print(print_plot)
# save figure
ggsave("../plot_new/fig_ndc_overall_four.pdf", plot = print_plot, 
       width = 12, height = 8, units = "in")

```{r ndc time stratified age strata, warning =F, echo = F, results='asis'} 
df_list <- list(beta_asthma, beta_copd, beta_both, beta_other)

outcome_list <- c("Asthma", "COPD", "Both", "Other")

# create an empty list to row bind dataframes together
datalist <- list()

# outcome name
outcome_name <- c("Asthma", "COPD", "Both", "Other")

# looking only at wrf chem, kriging, and geo-weighted; taking out 'Global Smoke' method
method_list <- c('WRF-Chem Smoke', 'Kriging Smoke', 'Geo-Weighted Smoke')

# create an empty list to row bind dataframes together
datalist1 <- list()
datalist2 <- list()

for(i in 1:length(df_list)){

  # dataframe to loop through
  df_to_loop <- data.frame(df_list[i])
  # indication of column
  outcome <- df_list[i]
  # outcome name
  outcome_name <- outcome_list[i]

  covariates_df <- df_to_loop[, c(1:26, 74:84)]

  # extract pm values and divide by 10 and ordered
  #which(colnames(df_to_loop)=="global_smk_pm_zip") # code to find column numbers
  pm_estimates_df <- df_to_loop[, c(87, 92, 91, 93)]/10  # create 10 unit increases

  # new loop for age categories
  for(k in 0:2){

    point_estimates <- matrix(nrow = 3, ncol = 10, byrow = T)

    # dataframe for analysis creation
    # bind columns back together 
    df_analysis <- cbind(covariates_df, pm_estimates_df) %>% 
      # remove missing pm values
      filter(!is.na(wrf_smk_pm_zip)) %>% 
      # limit to specific age category
      filter(age_ind == k) %>% 
      # the following code makes sure that the counterfactual values retained are 
      # symetric in that number of obs before = number of obs after
      mutate(obs_diff_admission = (fromdate - date)/7)

    for(j in 38:40){

    # combine previous values in dataframe that has all outcome/methods comparisons
    datalist1[[k+1]] <- point_est_df
  } # end age category loop

  # bind rows of age category estimates together
  age_est_df <- bind_rows(datalist1)

  # populate second dataframe list
  datalist2[[i]] <- age_est_df

} # end of outcome loop

# combine each outcome dataframe itteration in to a big dataset
combined_point_est_df <- bind_rows(datalist2)

  print(print_plot)
  # save figure
  ggsave("../plot_new/ndc_age_fig_four.pdf", plot = print_plot, 
       width = 12, height = 8, units = "in")

```{r ndc sex strata, warning = F, echo = F, results='asis'} 
df_list <- list(beta_asthma, beta_copd, beta_both, beta_other)

outcome_list <- c("Asthma", "COPD", "Both", "Other")

# outcome name
outcome_name <- c("Asthma", "COPD", "Both", "Other")

datalist1 <- list()
datalist2 <- list()

# data wrangling ----
# Producing conditional logit model estimates loop 
for(i in 1:length(df_list)){

  # dataframe to loop through
  df_to_loop <- data.frame(df_list[i])
  # indication of column
  outcome <- df_list[i]
  # outcome name
  outcome_name <- outcome_list[i]

  covariates_df <- df_to_loop[, c(1:26, 74:84)]

  # extract pm values and divide by 10 and ordered
  #which(colnames(df_to_loop)=="global_smk_pm_zip") # code to find column numbers
  pm_estimates_df <- df_to_loop[, c(87, 92, 91, 93)]/10  # create 10 unit increases

    for(j in 38:40){

  # bind rows of sex category estimates together
  sex_est_df <- bind_rows(datalist1)

  # populate second dataframe list
  datalist2[[i]] <- sex_est_df

} # end of outcome loop

# combine each outcome dataframe itteration in to a big dataset
combined_point_est_df <- bind_rows(datalist2)

  print(print_plot)
  # save figure
  ggsave("../plot_new/ndc_sex_fig_four.pdf", plot = print_plot, 
       width = 12, height = 8, units = "in")
